---
title: "Trust in Artificial Intelligent Agents Scale"
subtitle: "First quantitative approbation. Data preprocess workflow"
author:
  - name: Anton Angelgardt 
    affiliation: HSE University
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intro

Data was collected via [One Click Survey](https://1ka.si) platform. Participants were recruited via [Yandex.Toloka](https://toloka.yandex.ru) platform. Dataset contains Trust in Artificial Intelligent Agents Scale (TAIA) items, General Trust Scale (GT) items, questions aimed at TAIA scale validation and demographic data.


## Attributes description

### TAIA

Scale consists of six blocks. Each block contains 10--12 items.

* Predictability: `pr01`--`pr10`
* Consistency: `co01`--`co10`
* Utility: `ut01`--`ut12`
* Faith: `fa01`--`fa10`
* Dependability: `de01`--`de11`
* Understanding: `un01`--`un12`

Scores range from `0` to `5`. Reverse items are marked with `r` in its names.

Stems are [here](https://raw.githubusercontent.com/angelgardt/taia/master/preprocess/taia_items_1.txt).

Sequence of blocks was randomly shuffled for each participant as well as items sequence inside each block.

Questionnaire includes 15 auxiliary items `au01`--`au15` (2 or 3 per block) for controlling of questionnaire filling quality.



### GT

Questionnaire has one scale from six items `gt01`--`gt06`. Scores range from `0` to `4`. Scale has no reverse items.

Stems are [here](https://github.com/angelgardt/taia/raw/master/preprocess/gt_items.txt).



### Questions aimed at scale validation

Participants were asked to report their experience of dealing with AI technologies.

* `socnet` --- Does participant have an social network or social media accounts? (`1` for ‘yes‘, `0` for ‘no’)
  - if the answer is ‘no’, subject moves directly to `dighelp`
* social networks and social media accounts that participant have
  - `vk` for VKontakte
  - `fb` for Facebook
  - `tw` for Twitter
  - `in` for Instagram
  - `tt` for TikTok
  - `yt` for YouTube
* `freqvk`, `freqfb`, `freqtw`, `freqin`, `freqtt`, `freqyt` --- social media usage frequency (`0`--`3`)
* `expvk`, `expfb`, `exptw`, `expin`, `exptt`, `expyt` --- a level of satisfying with social media recommender system (`0`--`5`)
* `dighelp` --- Does participant use any digital assistants? (`1` for ‘yes‘, `0` for ‘no’)
  - if the answer is ‘no’, subject moves directly to `selfdrcar`
* digital assistants that participant use
  - `siri` for Siri (Apple)
  - `alice` for Alice (Yandex)
  - `salut` for Salut (Sberbank)
  - `oleg` for Oleg (Tinkoff Bank)
  - `alex` for Alexandra (Moscow Metro)
  - `mia` for Mia (Tele 2 mobile provider)
  - `mts` for MTS AI (MTS mobile provider)
  - `ggle` for Google Assistant
  - `oth` for ‘Other’ option
* `expsiri`, `expalice`, `expsalut`, `expoleg`, `expalex`, `expmia`, `expmts`, `expggle`, `expoth` --- How it was an interaction experience with these digital assistants? (`0`--`5`)
* `selfdrcar` --- Does participant have an experience of driving a self-driving car? (`1` for ‘yes‘, `0` for ‘no’)
  - if the answer is ‘no’, subject moves directly to `eduai`
* `selfdrexp` --- How it was an experience of driving a self-driving car? (`0`--`5`)
* `selfdrsafe` --- Did participant feel safe while driving a self-driving car? (`0`--`5`)
* `eduai` --- Does participant face with AI technologies in education? (`1` for ‘yes‘, `0` for ‘no’)
  - if the answer is ‘no’, subject moves directly to demographic part
* `eduaiexp` --- How it was an experience of dealing with AI for education? (`0`--`5`)



### Demographic info

* `age` --- participants’ age
* `sex` --- participants’ gender
  - `m` for ‘masculine’, `f` for ‘feminine’, `a` for ‘other’
* `edulvl1` --- a highest level of [first] education
  - `school` / `college` / `bachelor` / `specialist` / `master` / `postgrad`
* `spec1` --- [first] education specialization
* `edu2` --- Does participant have another education with different specialization? (`1` for ‘yes‘, `0` for ‘no’)
  - if the answer is ‘no’, subject moves directly to `jobfield`
* `edulvl2` --- a highest level of second education
  - `school` / `college` / `bachelor` / `specialist` / `master` / `postgrad`
* `spec2` --- second education specialization
* `jobfield` --- field of current job
* `joppos` --- job position
* `city` --- current city of residence


## Data preparation

Before preprocessing `id` column with unique string for each observation was add. Also extra characters added by One Click Survey platform were removed while personal data was separated from questionnaire data.





# Preprocessing

```{r echo=FALSE}
# load("preprocess.RData")
# save.image("preprocess.RData")
```

## Add packages

```{r}
library(tidyverse)
library(knitr)
```

## Import data

```{r}
raw <- read_csv('https://github.com/angelgardt/taia/raw/master/data/questionnaire_data.csv')
str(raw)
```


## Check quality with control items

Import keys for auxiliary items:

```{r}
aux_items <- read_csv2('https://github.com/angelgardt/taia/raw/master/data/aux_items.csv')
str(aux_items)
```

Select `au` columns from questionnaire data, merge with keys and calculate differences between score and key. Then calculate basic statistics for differences and assess quality for each participant.

```{r}
raw %>% select(id, all_of(aux_items$code)) %>% 
  pivot_longer(cols = aux_items$code,
               names_to = 'code',
               values_to = 'score') %>% 
  full_join(aux_items %>% 
              select(code, key)) %>% 
  mutate(dif = abs(score - key)) %>% 
  group_by(id) %>% 
  summarise(n_miss = sum(dif != 0),
            p_miss = n_miss / 15,
            min = min(dif),
            max = max(dif)) -> taia_quality
```

Quality of filling TAIA questionnaire assess based on control items:

```{r}
kable(taia_quality, 
      digits = 2,
      col.names = c("ID", "Number of misses", "Proportion of misses",
                    "Minimum miss", "Maximum miss")) %>% 
  head()
```

For excluding subject from analysis because of their inattentiveness a threshold of **33% misses (5 possible misses)** was chosen. Excluded participants are presented below.

```{r}
taia_quality %>% 
  filter(n_miss > 5) -> excluded_subjects
```

```{r}
kable(excluded_subjects,
      digits = 2,
      col.names = c("ID", "Number of misses", "Proportion of misses",
                    "Minimum miss", "Maximum miss"))
```

Total number of excluded subject is `r nrow(excluded_subjects)`.

Removing low quality observations:

```{r}
raw %>% 
  anti_join(excluded_subjects) -> taia
```

## Reverse items

Write function for reversinf items:

```{r}
reverse <- function(x) {
  as.numeric(
    dplyr::recode(x,
                "0" = "5",
                "1" = "4",
                "2" = "3",
                "3" = "2",
                "4" = "1",
                "5" = "0")
  )
}
```

Reverse items and remove control items. Rename reverse items:

```{r}
reversed_items <- str_extract(colnames(raw), pattern = regex("^[:alpha:]{2}[:digit:]{2}r$")) %>% na.omit()
```

```{r}
kable(reversed_items, col.names = "Reverse Items")
```
```{r}
taia %>%
  mutate_at(all_of(reversed_items), reverse) %>% 
  select(-all_of(aux_items$code)) %>% 
  rename_at(all_of(reversed_items),
            function(x) str_replace_all(x, "r$", "")) -> taia
```

```{r}
kable(head(taia))
```



# Integral score for General Trust Scale

```{r}
taia %>% 
  select(id, gt01, gt02, gt03, gt04, gt05, gt06) %>% 
  pivot_longer(cols = c("gt01", "gt02", "gt03", "gt04", "gt05", "gt06"),
               names_to = "gt_items", values_to = "gt_scores") %>%
  group_by(id) %>% 
  summarise(gt_score = mean(gt_scores)) %>% 
  full_join(taia) -> taia
```



# Fix data from questions

```{r}
taia %>% select(id, 74:118) %>% 
  mutate_all(function(x) ifelse(x < 0, NA, x)) -> questions
```

```{r}
questions %>% 
  select(id, vk, fb, tw, `in`, tt, yt) %>% 
  pivot_longer(c("vk", "fb", "tw", "in", "tt", "yt"),
               names_to = "socnets", values_to = "has_socnet") %>% 
  na.omit() %>% 
  group_by(id) %>% 
  summarise(n_socnet = sum(has_socnet)) %>% 
  full_join(taia) -> taia
```

```{r}
questions %>% 
  select(id, freqvk, freqfb, freqtw, freqin, freqtt, freqyt) %>% 
  pivot_longer(c("freqvk", "freqfb", "freqtw", "freqin", "freqtt", "freqyt"),
               names_to = "socnets", values_to = "freq_socnet") %>% 
  group_by(id) %>% 
  summarise(f_socnet = mean(freq_socnet, na.rm = TRUE)) %>% 
  mutate_at(vars(f_socnet), function(x) ifelse(is.nan(x), NA, x)) %>% 
  full_join(taia) -> taia
```

```{r}
questions %>% 
  select(id, expvk, expfb, exptw, expin, exptt, expyt) %>% 
  pivot_longer(c("expvk", "expfb", "exptw", "expin", "exptt", "expyt"),
               names_to = "socnets", values_to = "exp_socnet") %>% 
  group_by(id) %>% 
  summarise(e_socnet = mean(exp_socnet, na.rm = TRUE)) %>% 
  mutate_at(vars(e_socnet), function(x) ifelse(is.nan(x), NA, x)) %>% 
  full_join(taia) -> taia
```

```{r}
questions %>% 
  select(id, siri, alice, salut, oleg,
         alex, mia, mts, ggle, oth) %>% 
  pivot_longer(c("siri", "alice", "salut", "oleg",
                 "alex", "mia", "mts", "ggle", "oth"),
               names_to = "dighelper", values_to = "use_dighelper") %>% 
  na.omit() %>% 
  group_by(id) %>% 
  summarise(n_dighelp = sum(use_dighelper)) %>% 
  full_join(taia) -> taia
```

```{r}
questions %>% 
  select(id, expsiri, expalice, expsalut, expoleg,
         expalex, expmia, expmts, expggle, expoth) %>% 
  pivot_longer(c("expsiri", "expalice", "expsalut", "expoleg",
         "expalex", "expmia", "expmts", "expggle", "expoth"), 
         names_to = "dighelper", values_to = "exp_dighelper") %>% 
  group_by(id) %>% 
  summarise(e_dighelp = mean(exp_dighelper, na.rm = TRUE)) %>% 
  mutate_at(vars(e_dighelp), function(x) ifelse(is.nan(x), NA, x)) %>% 
  full_join(taia) -> taia
```

```{r}
taia %>% 
  mutate_at(vars(selfdrexp, selfdrsafe, eduaiexp),
            function(x) ifelse(x < 0, NA, x)) -> taia
```


# Export data

```{r}
str(taia)
```


```{r, eval=FALSE}
taia %>% 
  write_csv("~/taia/data/taia.csv")
```
